---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# surveyr

<!-- badges: start -->
[![R-CMD-check](https://github.com/danjdrennan/surveyr/workflows/R-CMD-check/badge.svg)](https://github.com/danjdrennan/surveyr/actions)
[![Codecov test coverage](https://codecov.io/gh/danjdrennan/surveyr/branch/main/graph/badge.svg)](https://app.codecov.io/gh/danjdrennan/surveyr?branch=main)
<!-- badges: end -->

The package is intended for use in an education setting working with survey data.
It provides tools for computing statistics from surveys using simple random
sampling and stratified designs.
Plotting is simplified with a thin wrapper on ggplot2 functions, enabling a user
to easily create aesthetic graphs.
Functions for deriving the sampling distribution of a small population and probability
demos are also included in the package. The probability demos cover the law of large
numbers and the central limit theorem.

## Installation

You can install the development version of surveyr from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("danjdrennan/surveyr")
```

## Example: Working with a Simple Random Sample

The most basic element for studies is a simple random sample with a known
population size. With data from a simple random sample, one can estimate the
population mean, total, or proportion using `mk_stat` as follows:

```{r srs}
# This library is designed to work alongside tidy packages,
# especially tidyr, tibble, dplyr, and ggplot2
library(dplyr)
library(tibble)
library(ggplot2)
library(surveyr)
# Generate a dataset to for README examples
set.seed(1)
d <- tibble::tibble(
    region = rep(1:5, 50),
    y = rnorm(250, 30 + 5*region, 5*sqrt(region))
)
N <- rpois(5, 100)

mk_stat(d$y, N=sum(N), stat="mean", fpc=TRUE)
```

The output gives a summary of the estimated population statistics, 
including the point estimate, variance of the estimate, standard error, and
coefficient of variation. Each column relates to the estimated statistic, not the
population. The `fpc` argument determines whether or not to use a finite population correction. In any case, the population size, `N`, must be supplied.


## Example: Working with a stratified dataset

If the data are grouped by some stratified variable (fixed effect), then a similar
data summary can be obtained using `make_summary`.

```{r stratified_data_table, echo=TRUE}
# Tabulate summary statistics for a stratified probability sample
make_summary(.data=d, .group=region, .y=y, .group_N=N, .fpc = TRUE, .stat="mean")
```

This table summary can be useful for comparing stratified groups, but it is often
an intermediate computation to estimating a total population statistic. Using
the pipe operator, we can compose functions to obtain that stratified result as
follows:

```{r data_summaries, echo=TRUE}
d %>% make_summary(.group=region, .y=y, .group_N=N, .fpc = TRUE, .stat="mean") %>%
    stratified_stat(.stat="mean", .fpc=TRUE)
```
Data plots can also be produced using `s_plot`, a thin wrapper on `ggplot2`.
The wrapper supports easy plotting of histograms and boxplots for simple random
samples or grouped data.

```{r plot_data, echo=TRUE}
d %>% s_plot(y, region, kind="hist")
```

# Example: Probability Demos

Demos of the Law of Large Numbers and the Central Limit Theorem are also provided.
The Law of Large Numbers draws samples from a particular Gamma distribution,
giving no flexibility to the user. The Central Limit Theorem demo actually
provides flexibility for choosing the distribution. In the CLT case, a user can
choose between a binormal distribution, a uniform distribution, or a gamma
distribution with any choice of parameterizations. Simply calling the functions
will generate plots using GGPlot as a backend, as can be seen next.

```{r lln, echo=TRUE}
lln <- lln_demo()
lln$plot
```

```{r clt, echo=TRUE}
clt <- clt_demo()
clt$plot
```

The central limit theorem demo also gives draws from the sampled distribution
for visualizing what the parent distribution looked like, along with the
theoretical parameters from the distribution, as can be seen below.

```{r clt_parent_dist, echo=TRUE}
clt$data %>% as_tibble %>% ggplot(aes(x=value)) +
    geom_histogram(bars=30, color="blue", fill="lightblue") + 
    geom_vline(xintercept = clt$popmean, size=1.2) +
    labs(
        title = "CLT Parent Distribution",
        subtitle = "Binormal Distribution with means 30, 50 and variance 36",
        x = "X",
        y = "Frequency"
    )
```
